### 2014-05-21 ITEPProceedings/piskunov.tex
Setup ITEP proceedings styling

\documentclass[
aps,%
12pt,%
final,%
notitlepage,%
oneside,%
onecolumn,%
nobibnotes,%
nofootinbib,% 
superscriptaddress,%
showpacs,%
centertags]%
{revtex4}

\newcommand{\exclude}[1]{}

\begin{document}
\selectlanguage{english}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

%\newcommand{\lll}{\lambda}

% Some useful journal names

\newcommand{\A}{\alpha}
\newcommand{\B}{\beta}
\newcommand{\T}{\theta}
\newcommand{\Ep}{\epsilon}
\newcommand{\beqn}{\begin{eqnarray}}
\newcommand{\eeqn}{\end{eqnarray}}

\title{ANOMALIES (SELECTED TOPICS)}

\author{\firstname{V.~I.}~\surname{Zakharov}}
\email{maxitg@icloud.com}
\affiliation{Institute for Theoretical and Experimental
Physics, Moscow, Russia
}

%\date{\today}

\begin {abstract}
The presentation is
intended  to be mainly pedagogical and covers the chiral anomaly
and
its most significant applications. Finally, we  
comment
on recent issues,
related to the lattice measurements.

\end{abstract}

\pacs{11.10.Ef,~11.15.Bt,~11.15.Ha}

\maketitle

\section{Introduction}
For the purposes of this review,
by anomalies one can understand violations
of the Ehrenfest theorem according to which
all matrix elements of the
classical equations of motion (considered as operators) vanish.
Thus, anomaly is established
 if we find, for example,
a nonvanishing
matrix element of the Dirac equation for a charged particle:
\begin{equation}
\langle \bar{\psi}(\hat{D}+im)\psi\rangle~\neq~0,
\end{equation}
where $\hat{D}=\gamma_{\mu}D_{\mu}$ and $D_{\mu}$ is the covariant derivative
and $m$
is the mass of the charged particle.

..........................

\bibliographystyle{maik}
\bibliography{piskunov}

\newpage

\selectlanguage{russian}

\begin{center}
\large \bfseries %\MakeTextUppercase{%
АНОМАЛИИ (ИЗБРАННЫЕ ТЕМЫ)
\end{center}
%
\begin{center}
{\bfseries
В.~И.~Захаров
}
\end{center}
%
\begin{center}
\begin{minipage}{\textwidth - 2cm}
\small

Данный обзор имеет образовательный характер и посвящен
киральным аномалиям и их наиболее существенным приложениям.
Также рассматриваются современные вопросы решеточных измерений.
\end{minipage}
\end{center}
\selectlanguage{english}
%
\newpage
{\bf Figure captions}
\vskip 1cm

{\bf Fig.~1.}~Anomalous triangle graph. Momenta $q,k_1,k_2$ are
carried by the axial current $a_\mu^5$ and
two photons, respectively. The triangle corresponds to a fermion of small mass $m$.

{\bf Fig.~2.}~Imaginary part of the anomalous triangle graph.
The intermediate fermions are on
mass shell. (The signs are the same as in Fig. 1)

{\bf Fig.~3.}~Three-body intermediate state in the imaginary part. (The signs are the same as in Fig. 1)

{\bf Fig.~4.}~Radiative correction to the two-body contribution to the
imaginary part. (The signs are the same as in Fig. 1)

{\bf Fig.~5.}~Graphic representation of the Banks--Casher evaluation of the quark condensate
$\langle \bar{q}q\rangle$. Dashed lines with crosses represent external (vacuum) gluonic
fields.

\newpage

%\begin{figure}
%\includegraphics{z1.eps}
%\caption{Zakharov, Fig. 1}
%\end{figure}

%\begin{figure}
%\includegraphics{z2.eps}
%\caption{Zakharov, Fig. 2}
%\end{figure}

%\begin{figure}
%\includegraphics{z3.eps}
%\caption{Zakharov, Fig. 3}
%\end{figure}

%\begin{figure}
%\includegraphics{z4.eps}
%\caption{Zakharov, Fig. 4}
%\end{figure}

%\begin{figure}
%\includegraphics{z5.eps}
%\caption{Zakharov, Fig. 5}
%\end{figure}

\end{document}
% \bibitem{kniel}
% Two-Loop Electroweak Corrections to the A0 gamma gamma and A0 g g Couplings of the CP-Odd Higgs Boson.
% Joachim Brod, Frank Fugel, Bernd A. Kniehl
%  Nucl.Phys.B807:188-209,2009.
% e-Print: arXiv:0807.1008 [hep-ph]